---
title: "Assignment 1, Task 1: Mountain Yellow-legged Frog"
author: "Claire Meuter"
date: "2023-01-26"
output: html_document
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, collapse = TRUE)
library(tidyverse)
library(readxl)
library(here)
library(lubridate)
library(janitor)
library(ggplot2)
library(patchwork)
```

## Overview of data 
This code explores, wrangles, and visualizes amphibian abundance data recorded by the Sierra Lakes Inventory Project. This data was collected from the Environmental Data Initiative (EDI) repository. As described from the Environmental Data Initiative repository: “The Sierra Lakes Inventory Project (SLIP) was a research endeavor that ran from 1995-2002 and has supported research and management of Sierra Nevada aquatic ecosystems and their terrestrial interfaces. We described the physical characteristics of and surveyed aquatic communities for >8,000 lentic water bodies in the southern Sierra Nevada, including lakes, ponds, marshes, and meadows.”

Metadata on the Sierra Lakes Inventory Project can be found here: https://portal.edirepository.org/nis/metadataviewer?packageid=edi.577.2



## Data exploration, wrangling, and graph building
Task 1: Creating a graph of mountain yellow-legged frog (Rana muscosa) count each year and across all water bodies, by life stage excluding the ‘EggMass’ level.
```{r}
#read in the data
sierra_amp <- read_excel(here("sierra_amphibians.xlsx"))

#view class of our date column
class(sierra_amp$survey_date)
#this tells me it's a numeric, but I need it in date format
sierra_amp$survey_date <- ymd(sierra_amp$survey_date)

#Now I check out the class again
class(sierra_amp$survey_date)
#Now it's in date, so we can use the year function

sierra_amp$survey_date <- year(sierra_amp$survey_date)

#Now sierra_amp only has year in survery_date column

#Now I'm going to clean the data to only include mountain yellow-legged frog (Rana muscosa) and exclude life stage  ‘EggMass’ level
#from reading the metadata I know Rana muscosa is code RAMU

rana_sierra <- sierra_amp %>%
  filter(amphibian_species == "RAMU") %>% #filtering for only RAMU
  filter(amphibian_life_stage != "EggMass") #dropping eggmass 
 

#use group by and summarise_at to find counts 
rana_count <- rana_sierra %>% 
  group_by(survey_date, amphibian_life_stage) %>% 
  summarise_at(vars(amphibian_number), #summing amphibian number column using survey_date and lifestage
               list(count=sum))

#Now I'll make a graph 
plot1 <- ggplot(rana_count, aes(fill=amphibian_life_stage, y= count, x=survey_date)) + scale_x_continuous(breaks = seq(from = 1995, to = 2002, by = 1)) + #making x axis annual with every year showing up 
    geom_bar(position="stack", stat="identity") + labs(y = "Count", x = "Survey Year", fill = "Life Stage") + theme_minimal() +
  theme(legend.position = c(0.2, 0.8))
```



Task 2: Creating a column graph containing total counts (over all years of the study) of combined adult and subadult endangered mountain yellow-legged frogs (Rana muscosa) observed in the 5 lakes with the greatest total observed counts.
```{r}
#First I need to do some data wrangling to get my data ready to graph. I can use the rana_sierra dataset. Then I need to exclude tadpoles from the data, make sure survey date is being read as categorical, and reformat if not. Next, I need to sum the amphibian_number by lake_id, and keep the 5 lakes that had to most frog counts. 

class(rana_sierra$lake_id)
#right now, class of lake_id is numeric, which I'll want to change

#making new dataset 
rana_frogs <- rana_sierra %>% 
  filter(amphibian_life_stage != "Tadpole") %>% # getting rid of tadpole so only subadult and adult are left 
  mutate( lake_id = as.character(lake_id)) %>% #change lake_id to a character class
  group_by(lake_id) %>%
  drop_na() %>% 
  summarize_at(vars(amphibian_number), #sum amphibian numbers within lake_id groups
               list(count = sum)) %>% 
    slice_max(order_by = count, n = 5) %>% #keep the 5 lakes with highest frog count 
    arrange(-count)    #sort the columns by amount of frogs
    
#renaming the lakes 
rana_frogs$lake_id[rana_frogs$lake_id == "50183"] <- "Lake 50183" 
rana_frogs$lake_id[rana_frogs$lake_id == "10223"] <- "Lake 10223"  
rana_frogs$lake_id[rana_frogs$lake_id == "10329"] <- "Lake 10329"  
rana_frogs$lake_id[rana_frogs$lake_id == "20196"] <- "Lake 20196"  
rana_frogs$lake_id[rana_frogs$lake_id == "10225"] <- "Lake 10225"

#graphing everything
plot2 <- ggplot(rana_frogs, aes(y=count, x = fct_reorder(lake_id, count))) + #fct_reorder allows me to sort the lake_id from least to greatest
  geom_col(fill="lightblue") +
  labs(x= "Lake ID", y = "Frog Count")+ theme_minimal()
  
```

## Compound figure
Graph of count of Mountain yellow-legged frogs by life stage combined with graph of total counts of combined adult and subadult  mountain yellow-legged frogs (Rana muscosa) observed in the 5 lakes with the greatest total observed count.
```{r, fig.height=7,fig.width=10}
plot1+plot2 #using patchwork to knit my graphs together
```
**Figure 1:** Counts of mountain yellow-legged frogs (*Rana mucosa*) recorded in the southern Sierra Nevada by the Sierra Lakes Inventory Project between 1995 and 2002. *Left:* Annual counts of *Rana mucosa*, differentiated by life stage using color codes. *Right:* Total number of *Rana mucosa* in the 5 lakes with the highest number of adult and subadult frogs observed. Source: Knapp et al. 2020.

## Citation

**Data citation:** Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3
